NLIZE: a Perturbation-Driven Visual Interrogation Tool for Analyzing and Interpreting Natural Language Inference Models

### 2018-03-26 paper/attention.tex

\begin{figure*}[t]
\centering
\vspace{-2mm}
\includegraphics[width=1.0\linewidth]{attentionView}
\caption{
Attention visualization.
}
\label{fig:attentionVis}
\end{figure*}

@@

As illustrated in Fig.~\ref{fig:attentionVis}, we employ too visual encoding for attention. In the grpah attention view (a), a bipartite graph encoding is adopted, in which the edge thickness corresponds to the attention value. Such a view is good for highlight the most dominate alignments. However, if many attention values are high, the edges may become cluttered lead to less effective visualization. In addition, if the sentence structure different between premise and hypothesis is drastically different, we are likely to see the prominent edges cross each other, which can also lead to confusing visual patterns. The matrix attention view (b), despite being more verbose and less efficient in highlighting the dominate alignment, do not have similar shortcomings. Together them complement each other and provide the same information from different perspective. As illustrated by the pink arrow line, we can see how the same attention value is visualization in both the grpah and the matrix view.
To help the user recongnize the correspondence during the exploration, we enable the linkage between highlight actions in both view.

>>>>

\begin{figure*}[t]
\centering
\vspace{-2mm}
\includegraphics[width=1.0\linewidth]{attentionView}
\vspace{-6mm}
\caption{
Attention visualization. In the graph attention view (a), a bipartite graph encoding is adopted, in which the edge thickness corresponds to the attention value. In the matrix attention view (b), the $i^{th}$ row corresponds to the probability of words in hypotheses align to the $i^{th}$ word in the premise.
The user can alter the attention values via the pop up interface illustrated in (c).
}
\label{fig:attentionVis}
\end{figure*}

@@

As illustrated in Fig.~\ref{fig:attentionVis}, we employ too visual encoding for attention. In the graph attention view (a), a bipartite graph encoding is adopted, in which the edge thickness corresponds to the attention value. Such a view is good for highlight the most dominate alignments. However, if many attention values are high, the edges may become cluttered lead to less effective visualization. In addition, if the sentence structure different between premise and hypothesis is drastically different, we are likely to see the prominent edges cross each other, which can also lead to confusing visual patterns. The matrix attention view (b), despite being more verbose and less efficient in highlighting the dominate alignment, do not have similar shortcomings. Together them complement each other and provide the same information from different perspective. As illustrated by the pink arrow line, we can see how the same attention value is visualization in both the grpah and the matrix view.
To help the user recongnize the correspondence during the exploration, we enable the linkage between highlight actions in both view.

### 2018-03-26 paper/overview.tex

@@

\section{NLIZE System}
In this section, we describe the design and implementation of the NLIZE (pronounced as ``analyze'') system, which address the analysis tasks (T1-3).
%
As discussed in the task analysis, the experts obtain intuition about the model via guided exploration by studying how various parts of the pipeline behave when certain components are altered.
%
This can be generalized as the perturbation-driven paradigm, which formulated the basic guiding principal for designing the proposed tool.
As illustrated in Fig.~\ref{fig:modelPipeline}, we enables the automated or user-guided perturbation of the input sentence (e.g., replace words), the attention in the model (e.g., alter the alignment between sentences), and the prediction (e.g., adjust the prediction by making updates to the model).
%
In the following sections, we describe in detail the five major components of the proposed system, namely, all pair summary view, sentence view, attention view, prediction view, pipeline view.

>>>>

\section{NLIZE System}
In this section, we discuss the design and implementation of the NLIZE (pronounced as ``analyze'', see interface overview in Fig.~\ref{fig:teaser}) system, which address the analysis tasks (\textbf{T1-3}).
%
As discussed in the task analysis, the experts often analyze and obtain intuition about the model by studying how altering one part of the model affect other stages of the pipeline.
%
Such a process can be generalized as the perturbation-driven paradigm, which is used as guiding principal for designing the proposed tool.
%
As illustrated in Fig.~\ref{fig:modelPipeline}, we enables the automated or user-guided perturbation of the input sentence (e.g., replace words), the attention in the model (e.g., alter the alignment between sentences), and the prediction (e.g., adjust the prediction by making updates to the model).
%
In the following sections, we describe in detail the five major components of the proposed system, namely, sentence view  (Section~\ref{sec:sentence}), attention view (Section~\ref{sec:attention}), prediction view (Section~\ref{sec:prediction}), pipeline view (Section~\ref{sec:pipeline}), all pair summary view (Section~\ref{sec:allPairs}).

@@

### 2018-03-26 paper/prediction.tex
add placeholder test
update teaser

\subsection{User-Guided Prediction Update}

\subsubsection{Margin-Infused Relaxed Algorithm (MIRA)}
the mira loss is $|w' - w|^2 + C * J(w')$

>>>>

\subsection{Prediction View}
\label{sec:prediction}
For a given sentence pair, the model predicts a discrete probability distribution of the three labels (neutral, contradict, and entailment).
%
In the prediction view, as illustrated in Fig.~\ref{fig:predictionView}, a prediction probability is encoded as a point in the barycentric coordinate system of the triangle. In the triangle, the distinct color indicates the region corresponds to different labels. The prediction result for the original sentence pair is represented by the larger yellow circle, whereas smaller grey circles illustrate the perturbed sentence pairs. A density contour of the prediction is computed to emphasis the highly cluttered areas and distinguish the outliers.

To address \textbf{T3}, we need a way to communicate the reassignment of the prediction. As illustrated in (b), we integrate such operation in the prediction panel. When press and drag the circle (prediction), the user is presented with the three option (E, N, C) corresponds to the labels. When the user hovers on one of the options, a dotted line is shown to indicate the newly assigned label.

\begin{figure}[htbp]
\centering
\vspace{-2mm}
 \includegraphics[width=1.0\linewidth]{predictionView}
 \vspace{-2mm}
 \caption{
In the prediction view, the prediction is encoded as a point in the barycentric coordinate system of the triangle shown in (a).
%
The prediction result for the original sentence pair is represented by a larger yellow circle and the prediction of perturbed pairs are illustrated by smaller grey circles.
A density contour of the prediction is computed to emphasize the highly cluttered areas and distinguish the outliers.
As illustrated in (b), the predicted label does not match the ground truth, therefore, we apply a label reassignment operation, which triggers the model update.
% when press and drag the circle (prediction), the user is presented with the three option corresponds to the labels. The dotted line and arrow indicate the newly assigned label.
 }
\label{fig:predictionView}
\end{figure}

### 2018-03-26 paper/sentence.tex

\subsection{Sentence View}
To start an exploration cycle, the experts need to examine the premise and hypothesis sentence the current example. 
As illustrated in Fig.~\ref{fig:modelPipeline}, the sentence view shows the premise and hypothesis sentence in (c) and (d).
%
To facilitate the analysis task (\textbf{T1}), we employ an automate sentence perturbation scheme that replaces nous and verb by their synonymous in the wordNet~\cite{Miller1995} (the standard lexical database for NLP applications).
%
When perturbation is applied to either of the sentences, the replaced words are highlight in blue (see Fig~\ref{fig:modelPipeline}(e) ) to indicate the modification to the original sentence.
%
In (a), \emph{Predict} and \emph{Predict All}, each corresponds to predict the current displayed sentence pairs, or predict all combinations of perturbed premise and hypothesis, provided there is only one perturbation per sentence pair (i.e., we use original premise if hypothesis is perturbed, or use original hypothesis if premise is perturbed). 
%%% sentence perturbation %%%%
In (b), previous explored original sentence are stored in the dropdown list that allow the user to revisit previous examined sentence.
In addition, the user can type any sentences or modified existing text in the (c) and (d) to accommodate customized input when expert want to conduct quick experiment or when the automatic perturbation does not meet expectations.

\begin{figure}[htbp]
\centering
\vspace{-2mm}
 \includegraphics[width=0.95\linewidth]{sentenceView}
 \caption{
Sentence view shows the premise (c) and the hypothesis (d) sentence pair. The word that replaces the original text is highlight in blue, as illustrated in (e)
In (a), \emph{Predict} and \emph{Predict All}, each corresponds to predict the current displayed sentence pairs, or predict all combinations of perturbed premise and hypothesis.
In (b), previous explored original sentence are stored in the dropdown list that allow the user to revisit previous examined sentence.
 }
\label{fig:modelPipeline}
\end{figure}

>>>>

\subsection{Sentence View}
\label{sec:sentence}
To start an exploration cycle, the experts need to examine the premise and hypothesis sentence the current example.
As illustrated in Fig.~\ref{fig:sentenceView}, the sentence view shows the premise and hypothesis sentence in (c) and (d).
%
To facilitate the analysis task (\textbf{T1}), we employ an automate sentence perturbation scheme that replaces nous and verb by their synonymous in the wordNet~\cite{Miller1995} (the standard lexical database for NLP applications).
%
When perturbation is applied to either of the sentences, the replaced words are highlight in blue (see Fig~\ref{fig:modelPipeline}(e) ) to indicate the modification to the original sentence.
%
In (a), \emph{Predict} and \emph{Predict All}, each corresponds to predict the current displayed sentence pairs, or predict all combinations of perturbed premise and hypothesis, provided there is only one perturbation per sentence pair (i.e., we use original premise if hypothesis is perturbed, or use original hypothesis if premise is perturbed).
%%% sentence perturbation %%%%
In (b), previous explored original sentence are stored in the dropdown list that allow the user to revisit previous examined sentence.
In addition, the user can type any sentences or modified existing text in the (c) and (d) to accommodate customized input when expert want to conduct quick experiment or when the automatic perturbation does not meet expectations.

\begin{figure}[htbp]
\centering
\vspace{-2mm}
 \includegraphics[width=0.95\linewidth]{sentenceView}
 \caption{
Sentence view shows the premise (c) and the hypothesis (d) sentence pair. The word that replaces the original text is highlight in blue, as illustrated in (e)
In (a), \emph{Predict} and \emph{Predict All}, each corresponds to predict the current displayed sentence pairs, or predict all combinations of perturbed premise and hypothesis.
In (b), previous explored original sentence are stored in the dropdown list that allow the user to revisit previous examined sentence.
 }
\label{fig:sentenceView}
\end{figure}
